The Revisited Hidden Weight Bit Function

### 2024-12-13 RevisedHiddenWeight.tex
minor

\begin{definition}[Balancedness]
%    A Boolean function $f\in\BN$ is called balanced if $\wt f(0_n)=0$.
A Boolean function $f\in\BN$ is called balanced if $|\{x \in \F_2^n \mid f(x)=0\}|=|\{x \in \F_2^n \mid f(x)=1\}|=2^{n-1}$. 
Equivalently, $f$ is balanced if and only if $\wt f(0_n)=0$.
\end{definition}

@@

\begin{proof}
	We use Proposition~\ref{prop:WT}, applying Remark~\ref{rmk:dependence_dnka} and that $\pi^{-1}(0_n)=0_n$:
	\begin{align*}
		\wt{f}(0_n)&=1+\sum_{k=1}^n\Dnka{n}{k}{\pi^{-1}(0_n)+(0,1,0,1,\ldots,0,1)+\pi(e_k)}\\
		&=1+\sum_{k=1}^{n/2}\Dnka{n}{k}{1,1,0,1,\ldots,0,1}+\sum_{k=n/2+1}^{n}\Dnka{n}{k}{0,0,0,1,\ldots,0,1}.
	\end{align*}
	Using the recursive relation from Proposition~\ref{prop:recursiveDnka}, we obtain:
	\[
		\Dnka{n}{k}{1,1,0,1,\ldots,0,1}=\Dnka{n-2}{k}{1,1,0,1,\ldots,0,1} +\Dnka{n-2}{k-2}{1,1,0,1,\ldots,0,1}.
	\]
	By reapplying this $n/2-1$ times, we obtain:
	\begin{align*}
		\Dnka{n}{k}{1,1,0,1,\ldots,0,1}&=\sum_{i=0}^{n/2-1} \binom{n/2-1}{i} \Dnka{2}{k-2i}{1,1}\\
		&=\binom{n/2-1}{k/2} \Dnka{2}{0}{1,1}+ \binom{n/2-1}{(k-1)/2} \Dnka{2}{1}{1,1}\\
		&\qquad+ \binom{n/2-1}{k/2-1} \Dnka{2}{2}{1,1}\\
		&=\binom{n/2-1}{k/2}-2\binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}.
	\end{align*}
	We similarly get $\Dnka{n}{k}{0,0,0,1,\ldots,0,1}=\binom{n/2-1}{k/2} +2 \binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}$. Then, combining both, we obtain:
	\begin{align*}
		\wt{f}(0_n)&=1+\sum_{k=1}^{n/2}\Dnka{n}{k}{1,1,0,1,\ldots,0,1}+ \sum_{k=n/2+1}^{n}\Dnka{n}{k}{0,0,0,1,\ldots,0,1}\\
		&=\sum_{k=0}^{n/2} \left(\binom{n/2-1}{k/2} -2 \binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}\right)\\
		&\qquad+\sum_{k=n/2+1}^{n}\left(\binom{n/2-1}{k/2}+2\binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}\right)\\
		&=\sum_{k=0}^{n}\left( \binom{n/2-1}{k/2}-\binom{n/2-1}{k/2-1}\right)+2\left(\sum_{k=n/2+1}^{n}\binom{n/2-1}{(k-1)/2}-\sum_{k=0}^{n/2}\binom{n/2-1}{(k-1)/2}\right)\\
		&=0+2\left(\sum_{k=n/2+1}^{n}\binom{n/2-1}{(k-1)/2}-\sum_{k=0}^{n/2} \binom{n/2-1}{(k-1)/2}\right)\\
		&=2\left(\sum_{i=\lceil n/4\rceil}^{n/2-1} \binom {n/2-1}{i} - \sum_{i=0}^{\lfloor(n-2)/4\rfloor} \binom{n/2-1}{i}\right).
	\end{align*}
	The announced result follows by applying that $\binom{m}{k}=\binom{m}{m-k}$ for all integers $m\geq 0$ and $k$.
\end{proof}

@@

\begin{proof}
    It is enough to use the expressions from Table \ref{tab:Dnka_small_k}, and replace $p$, $q$ and $r$ by their respective values. For instance, for the case $n=16m$, we replace $p$, $q$ and $r$ by $m+s(s-1)$, $m+s(s+1)$ and $6m-2s^2$ in the expression $\Dnka n2a=2(p-q)^2-3(p+q)+r$, which yields $\Dnka n2a=0$. This checking can be autotomized by a computer algebra system.
	\ifnum\anon=0
		For this, we refer to the corresponding \textsf{SageMath} implementation, accessible via this link: \url{https://github.com/se-tim/Revisited-HWBF.git}.
	\fi
\end{proof}

@@

\begin{proof}
	For the sum and the alternating sum, it is enough to compute $P_a(1)$ and $P_a(-1)$, respectively. The sum of the $\Dnka nka$ over the even (respectively, odd) $k$ is obtained by adding (respectively, subtracting) these two sums and dividing by $2$.
\end{proof}

@@

In this part, we generalize the results of Section~\ref{sec:Cauchy} to all weightwise quadratic functions $f$ such that for every $k\in [1,n]$, the function $f_k$ in Definition~\ref{def:wwdegd} has exactly $t$ quadratic terms with no shared variables.

@@

    Similarly to the proof of Theorem~\ref{theorem:bound_D_nka}, we want to bound the maximum of $|P_{t,a}(z)|$ on $\dis$ for $a\in\F_2^n$, and this will then be an upper bound on $|\Dtkna tkna|$ for every $k\in[0,n]$. 
    We write $a=(b,c)$ with $b\in\F_2^{2t}$ and $c\in\F_2^{n-2t}$, and define $(p,q,r)=(p(b),q(b),r(b))$, and further consider $u=n-2t-\w(c)$ and $v=\w(c)$ as in Definition~\ref{defi:P_a_t}.    
	To find a bound for $|P_{t,a}(z)|$ on $\dis$ that applies to all $a\in\F_2^n$, we would like to bound the quantity $\max_{a\in\F_2^n,z\in\dis}|P_{t,a}(z)|$. Moving through all $a\in\F_2^n$ is equivalent to moving through all tuples $(p,q,r;u,v)$ of non-negative integers which satisfy $p+q+r=t$ and $u+v=n-2t$. It follows from Lemma~\ref{lemma:maximizing_args} that it is enough to consider only the cases for which only one of $p,q,r$ is non-zero, and for which only one of $u,v$ is non-zero. 
	This leaves us with the following cases to consider for $(p,q,r;u,v)$:
    \begin{gather*}
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0),\\
        (t,0,0;0,n-2t),\quad(0,t,0;0,n-2t),\quad(0,0,t;0,n-2t).
    \end{gather*}
    Since simultaneously replacing $(p,q)$ by $(q,p)$ and $(u,v)$ by $(v,u)$ has the same effect as replacing $z$ by $-z$ in $P_{t,a}(z)$ and therefore does not change the maximum of $|P_{t,a}(z)|$ on $\dis$, we can further restrict ourselves to the following tuples for $(p,q,r;u,v)$:
    \[
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0).
    \]

@@

    Therefore, in case $\lambda<1/6$, we can conclude from the two claims that $g(\alpha)$ reaches its maximum at $\alpha_0=0$, and this maximum is then equal to $2^{2-2\lambda}=2^{2\mu}$, as required. For $\lambda=1/6$, the solution to the equation $\mathcal E$ is $\cos\alpha=1$, again implying that $g(\alpha)$ reaches its maximum at $\alpha_0=0$, and we obtain the same maximum of $2^{2\mu}$. In case $\lambda>1/6$, the two claims imply that $g(\alpha)$ reaches its maximum at the unique $\alpha_0\in[0,\pi]$ for which $\cos\alpha_0=-\lambda+\sqrt{\lambda^2-4\lambda+2}$. For this $\alpha_0$, the double angle formula for the cosine implies that $\cos(2\alpha_0)=4\lambda^2-8\lambda+3-4\lambda\sqrt{\lambda^2-4\lambda+2}$. If we replace $\cos\alpha_0$ and $\cos(2\alpha_0)$ by their respective values, we obtain that the maximum of $g(\alpha)$ for $\alpha\in[0,\pi]$ is given by the following value:
    \[
        g(\alpha_0)=\frac{\left(-8\lambda^2+16\lambda+8\lambda\sqrt{\lambda^2-4\lambda+2}\right)^{\lambda}}{\left(2-2\lambda+2\sqrt{\lambda^2-4\lambda+2}\right)^{2\lambda-1}}=2^{2\mu}.
    \]
    With this, we have covered the case $(p,q,r;u,v)=(t,0,0;n-2t,0)$ entirely. Similarly, it can be proven that $|P_{t,a}(z)|$ is also bounded by $2^{\mu n}$ on $\dis$ for the remaining two cases of $(p,q,r;u,v)$. In fact, for $(p,q,r;u,v)=(0,t,0;n-2t,0)$, one can proceed in exactly the same way, and for $(p,q,r;u,v)=(0,0,t;n-2t,0)$, it is enough to replace $g(\alpha)$ by $h(\alpha)=(2+2\cos(2\alpha))^\lambda(2+2\cos\alpha)^{1-2\lambda}$, whose maximum is bounded by the maximum of $g(\alpha)$ because $0\leq h(\alpha)\leq g(\alpha)$.
\end{proof}

@@

\begin{proof}
	Again, let $q\in\BN$ be given by $q(x)=\sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}$, so that $h=f+q$.
	Let $g\in\BN^*$ with $\degg(g)=\AI(f)$ be an annihilator of $f+ \varepsilon$ for some $\varepsilon\in\{0,1\}$. Then the following holds:
	\[g\cdot(q+1) \cdot(h+\varepsilon) = g\cdot(q+1)\cdot(f+q+\varepsilon)=g\cdot(q+1)\cdot(f+\epsilon)+g\cdot(q+1)\cdot q=0+0=0.\]
	Hence, $g\cdot(q+1)$ is an annihilator of $h+ \varepsilon$. 
	If $g\cdot(q+1)\ne 0$, it follows that $\AI(h)\leq\degg(g\cdot(q+1))\leq\degg(g)+\degg(q+1)$, and therefore that $\AI(f)=\degg(g)\geq\AI(h)-\degg(q+1)=\AI(h)-2$. If $g\cdot(q+1)=0$, then $0=g\cdot(f+\epsilon+q+1)=g\cdot(h+\epsilon+1)$, showing that $\AI(f)=\degg(g)\geq\AI(h)$.
\end{proof}

>>>>

\begin{definition}[Balancedness]
%    A Boolean function $f\in\BN$ is called balanced if $\wt f(0_n)=0$.
A Boolean function $f\in\BN$ is called balanced if $|\{x \in \F_2^n : f(x)=0\}|=|\{x \in \F_2^n : f(x)=1\}|=2^{n-1}$. 
Equivalently, $f$ is balanced if and only if $\wt f(0_n)=0$.
\end{definition}

@@

\begin{proof}
	We use Proposition~\ref{prop:WT}, applying Remark~\ref{rmk:dependence_dnka} and that $\pi^{-1}(0_n)=0_n$:
	\begin{align*}
		\wt{f}(0_n)&=1+\sum_{k=1}^n\Dnka{n}{k}{\pi^{-1}(0_n)+(0,1,0,1,\ldots,0,1)+\pi(e_k)}\\
		&=1+\sum_{k=1}^{n/2}\Dnka{n}{k}{1,1,0,1,\ldots,0,1}+\sum_{k=n/2+1}^{n}\Dnka{n}{k}{0,0,0,1,\ldots,0,1}.
	\end{align*}
	Using the recursive relation from Proposition~\ref{prop:recursiveDnka}, we obtain:
	\[
		\Dnka{n}{k}{1,1,0,1,\ldots,0,1}=\Dnka{n-2}{k}{1,1,0,1,\ldots,0,1} +\Dnka{n-2}{k-2}{1,1,0,1,\ldots,0,1}.
	\]
	By reapplying this $n/2-1$ times, we obtain:
%	\begin{align*}
%		\Dnka{n}{k}{1,1,0,1,\ldots,0,1}&=\sum_{i=0}^{n/2-1} \binom{n/2-1}{i} \Dnka{2}{k-2i}{1,1}\\
%		&=\binom{n/2-1}{k/2} \Dnka{2}{0}{1,1}+ \binom{n/2-1}{(k-1)/2} \Dnka{2}{1}{1,1}\\
%		&\qquad+ \binom{n/2-1}{k/2-1} \Dnka{2}{2}{1,1}\\
%		&=\binom{n/2-1}{k/2}-2\binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}.
%	\end{align*}
	\begin{align*}
\Dnka{n}{k}{1,1,0,1,\ldots,0,1}&=\sum_{i=0}^{n/2-1} \binom{n/2-1}{i} \Dnka{2}{k-2i}{1,1}\\
&=\binom{n/2-1}{k/2} \Dnka{2}{0}{1,1}+ \binom{n/2-1}{(k-1)/2} \Dnka{2}{1}{1,1}+ \binom{n/2-1}{k/2-1} \Dnka{2}{2}{1,1}\\
&=\binom{n/2-1}{k/2}-2\binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}.
\end{align*}
We similarly ge
	We similarly get $\Dnka{n}{k}{0,0,0,1,\ldots,0,1}=\binom{n/2-1}{k/2} +2 \binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}$. Then, combining both, we obtain:
	\begin{align*}
		\wt{f}(0_n)&=1+\sum_{k=1}^{n/2}\Dnka{n}{k}{1,1,0,1,\ldots,0,1}+ \sum_{k=n/2+1}^{n}\Dnka{n}{k}{0,0,0,1,\ldots,0,1}\\
		&=\sum_{k=0}^{n/2} \left(\binom{n/2-1}{k/2} -2 \binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}\right)\\
		&\qquad+\sum_{k=n/2+1}^{n}\left(\binom{n/2-1}{k/2}+2\binom{n/2-1}{(k-1)/2}-\binom{n/2-1}{k/2-1}\right)\\
		&=\sum_{k=0}^{n}\left( \binom{n/2-1}{k/2}-\binom{n/2-1}{k/2-1}\right)+2\left(\sum_{k=n/2+1}^{n}\binom{n/2-1}{(k-1)/2}-\sum_{k=0}^{n/2}\binom{n/2-1}{(k-1)/2}\right)\\
		&=0+2\left(\sum_{k=n/2+1}^{n}\binom{n/2-1}{(k-1)/2}-\sum_{k=0}^{n/2} \binom{n/2-1}{(k-1)/2}\right)\\
		&=2\left(\sum_{i=\lceil n/4\rceil}^{n/2-1} \binom {n/2-1}{i} - \sum_{i=0}^{\lfloor(n-2)/4\rfloor} \binom{n/2-1}{i}\right).
	\end{align*}
	The final result follows by applying that $\binom{n}{k}=\binom{n}{n-k}$ for all integers $n\geq 0$ and $k$.
\end{proof}

@@

\begin{proof}
 %   It is enough to use the expressions from Table \ref{tab:Dnka_small_k}, and replace $p$, $q$ and $r$ by their respective values. For instance, for the case $n=16m$, we replace $p$, $q$ and $r$ by $m+s(s-1)$, $m+s(s+1)$ and $6m-2s^2$ in the expression $\Dnka n2a=2(p-q)^2-3(p+q)+r$, which yields $\Dnka n2a=0$. This checking can be autotomized by a computer algebra system.
 We use the expressions from Table~\ref{tab:Dnka_small_k}, and replace $p$, $q$ and $r$ by their respective values. 
 For instance, for the case $n=16m$, we replace $p$, $q$ and $r$ by $m+s(s-1)$, $m+s(s+1)$ and $6m-2s^2$ in the expression $\Dnka n2a=2(p-q)^2-3(p+q)+r$, which yields $\Dnka n2a=0$. ~\footnote{This checking can be automatized by a computer algebra system.
	\ifnum\anon=0
		For this, we refer to the corresponding \textsf{SageMath} implementation, accessible via this link: \url{https://github.com/se-tim/Revisited-HWBF.git}.
	\fi
}
\end{proof}

@@

\begin{proof}
	For the sum and the alternating sum, we compute $P_a(1)$ and $P_a(-1)$, respectively. 
	The sum of the $\Dnka nka$ over the even (respectively, odd) $k$ is obtained by adding (respectively, subtracting) these two sums and dividing by $2$.
\end{proof}

@@

%In this part, we generalize the results of Section~\ref{sec:Cauchy} to all weightwise quadratic functions $f$ such that for every $k\in [1,n]$, the function $f_k$ in Definition~\ref{def:wwdegd} has exactly $t$ quadratic terms with no shared variables.

In this part, we generalize the results of Section~\ref{sec:Cauchy} to all weightwise quadratic functions $f$ such that, for every $k \in [1, n]$, the function $f_k$ (as defined in Definition~\ref{def:wwdegd}) contains exactly $t$ quadratic terms with no shared variables.

@@

    Similarly to the proof of Theorem~\ref{theorem:bound_D_nka}, we want to bound the maximum of $|P_{t,a}(z)|$ on $\dis=\{ z \in \C: |z|=1  \}$ for $a\in\F_2^n$, and this will then be an upper bound on $|\Dtkna tkna|$ for every $k\in[0,n]$. 
    We write $a=(b,c)$ with $b\in\F_2^{2t}$ and $c\in\F_2^{n-2t}$, and define $(p,q,r)=(p(b),q(b),r(b))$, and further consider $u=n-2t-\w(c)$ and $v=\w(c)$ as in Definition~\ref{defi:P_a_t}.    
	To find a bound for $|P_{t,a}(z)|$ on $\dis$ that applies to all $a\in\F_2^n$, we would like to bound the quantity $\max_{a\in\F_2^n,z\in\dis}|P_{t,a}(z)|$. Moving through all $a\in\F_2^n$ is equivalent to moving through all tuples $(p,q,r;u,v)$ of non-negative integers which satisfy $p+q+r=t$ and $u+v=n-2t$. It follows from Lemma~\ref{lemma:maximizing_args} that it is enough to consider only the cases for which only one of $p,q,r$ is non-zero, and for which only one of $u,v$ is non-zero. 
	This leaves us with the following cases to consider for $(p,q,r;u,v)$:
    \begin{gather*}
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0),\\
        (t,0,0;0,n-2t),\quad(0,t,0;0,n-2t),\quad(0,0,t;0,n-2t).
    \end{gather*}
    Since simultaneously replacing $(p,q)$ by $(q,p)$ and $(u,v)$ by $(v,u)$ has the same effect as replacing $z$ by $-z$ in $P_{t,a}(z)$ and therefore does not change the maximum of $|P_{t,a}(z)|$ on $\dis$, we can further restrict ourselves to the following tuples for $(p,q,r;u,v)$:
    \[
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0).
    \]

@@

    Therefore, in case $\lambda<1/6$, we can conclude from the two claims that $g(\alpha)$ reaches its maximum at $\alpha_0=0$, and this maximum is then equal to $2^{2-2\lambda}=2^{2\mu}$, as required. For $\lambda=1/6$, the solution to the equation $\mathcal E$ is $\cos\alpha=1$, again implying that $g(\alpha)$ reaches its maximum at $\alpha_0=0$, and we obtain the same maximum of $2^{2\mu}$. In case $\lambda>1/6$, the two claims imply that $g(\alpha)$ reaches its maximum at the unique $\alpha_0\in[0,\pi]$ for which $\cos\alpha_0=-\lambda+\sqrt{\lambda^2-4\lambda+2}$. For this $\alpha_0$, the double angle formula for the cosine implies that $\cos(2\alpha_0)=4\lambda^2-8\lambda+3-4\lambda\sqrt{\lambda^2-4\lambda+2}$. If we replace $\cos\alpha_0$ and $\cos(2\alpha_0)$ by their respective values, we obtain that the maximum of $g(\alpha)$ for $\alpha\in[0,\pi]$ is given by the following value:
    \[
        g(\alpha_0)=\frac{\left(-8\lambda^2+16\lambda+8\lambda\sqrt{\lambda^2-4\lambda+2}\right)^{\lambda}}{\left(2-2\lambda+2\sqrt{\lambda^2-4\lambda+2}\right)^{2\lambda-1}}=2^{2\mu}.
    \]
    With this, we have covered the case $(p,q,r;u,v)=(t,0,0;n-2t,0)$ entirely. Similarly, it can be proven that $|P_{t,a}(z)|$ is also bounded by $2^{\mu n}$ on $\dis$ for the remaining two cases of $(p,q,r;u,v)$. 
    For $(p,q,r;u,v)=(0,t,0;n-2t,0)$, one can proceed in exactly the same way, and for $(p,q,r;u,v)=(0,0,t;n-2t,0)$, it is enough to replace $g(\alpha)$ by $h(\alpha)=(2+2\cos(2\alpha))^\lambda(2+2\cos\alpha)^{1-2\lambda}$, whose maximum is bounded by the maximum of $g(\alpha)$ since $0\leq h(\alpha)\leq g(\alpha)$.
\end{proof}

@@

\begin{proof}
	Let $q\in\BN$ be given by $q(x)=\sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}$, so that $h=f+q$.
	Let $g\in\BN^*$ with $\degg(g)=\AI(f)$ be an annihilator of $f+ \varepsilon$ for some $\varepsilon\in\{0,1\}$. 
	Then the following holds:
	\[g\cdot(q+1) \cdot(h+\varepsilon) = g\cdot(q+1)\cdot(f+q+\varepsilon)=g\cdot(q+1)\cdot(f+\epsilon)+g\cdot(q+1)\cdot q=0+0=0.\]
	Hence, $g\cdot(q+1)$ is an annihilator of $h+ \varepsilon$. 
	If $g\cdot(q+1)\ne 0$, it follows that $\AI(h)\leq\degg(g\cdot(q+1))\leq\degg(g)+\degg(q+1)$, and therefore that $\AI(f)=\degg(g)\geq\AI(h)-\degg(q+1)=\AI(h)-2$. If $g\cdot(q+1)=0$, then $0=g\cdot(f+\epsilon+q+1)=g\cdot(h+\epsilon+1)$, showing that $\AI(f)=\degg(g)\geq\AI(h)$.
\end{proof}
